---
title: "Lab06"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab06}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab06)
```
This vignette contains the report for lab 6: writing fast R code. In this lab, different algorithms with different complexity have been implemented for the Knapsack problem to study the effects of the different algorithms with a focus on speed.


Generate knapsack objects with n=2000 and n=1000000 items respectively:
```{r}
RNGversion(min(as.character(getRversion()),"3.5.3"))

set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")

n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000)
  )

n <- 100000
knapsack_objects_greedy <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000)
  )

```
We run the system.time() function 10 or 100 times with a for loop to obtain the average time of the function to reduce the time error and have a more precise comparison.

## 1.1.2 Brute force search

Time for n=16 objects:

```{r}
n_iter <- 1
n_select <- 16
system.time(for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500))/n_iter

system.time(for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 2000))/n_iter

```

It takes around 0.5s for both weights.

## 1.1.3 Dynamic programming

Time for n=500 objects:

```{r}
n_iter <- 1
n_select <- 500

system.time(for (i in 1:n_iter) knapsack_dynamic(x = knapsack_objects[1:n_select,],
                                                 W = 3500))/n_iter

system.time(for (i in 1:n_iter) knapsack_dynamic(x = knapsack_objects[1:n_select,], 
                                                 W = 2000))/n_iter

microbenchmark::microbenchmark(knapsack_dynamic(x = knapsack_objects[1:n_select,], 
                                                W = 3500),
                               knapsack_dynamic(x = knapsack_objects[1:n_select,], 
                                                W = 2000))

```
It takes around 0.7s for W=3500 and 0.3s for W=2000. It seems reasonable that the 
smaller weight takes shorter time since the table the algorithm has to 
construct becomes smaller with a smaller value for W.

## 1.1.4 Greedy heuristic

```{r}
n<-1
system.time(for (i in 1:n) greedy_knapsack(knapsack_objects[1:1200,],W=3500)) / n
#100000 Objects
system.time(for (i in 1:n)greedy_knapsack(knapsack_objects_greedy[1:100000,],
                                          W=3500))/n
```
How much time does it takes to run the algorithm for n = 1000000 objects?
It takes 36.018 seconds.
Approximately 35 seconds more than with 1200 objects.

## 1.1.5 Improvement of greedy heuristic

```{r}
n<-1
system.time(for (i in 1:n) greedy_knapsack(knapsack_objects[1:1200,],W=3500)) / n
greedy_knapsack(knapsack_objects[1:1200,],W=3500)
system.time(for (i in 1:n) greedy_knapsack(knapsack_objects[1:1200,],W=3500,"plus")) / n
greedy_knapsack(knapsack_objects[1:1200,],W=3500,"plus")
```
"Plus" optimization, finds the next item which Weight value fits the  Weight 
limit of the knapsack. In this particular case is the item 764, as we can see 
above.

## 1.1.6 Profile your code and optimize your code

We compared for different main approaches to do the independent steps of the 
brute force algorithm, and within two of the main approaches also some sub types. 
The approaches were:

- for loop
- matrix calculations
- apply
    1a) part of the steps used variants of apply, and parts used matrix algebra
    1b) all steps used variants of apply, in separate calls
    2) all steps were collected into one function, that was used in apply (sapply)

- parallel package
    1a) part of the steps used variants of parApply, and parts used matrix algebra
    1b) all steps used variants of parApply, in separate calls
    2) all steps were collected into one function, that was used in parApply (parSapply)

The comparisons were made using W=3500.    
```{r}
n_iter <- 1
n_select <- 18

#- For loop
brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="forloop")
system.time(for (i in 1:n_iter) brute_force_knapsack(
  x = knapsack_objects[1:n_select,], W = 3500, type="forloop")) /n_iter

#- Matrix --------------------#
brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="matrix")
system.time(for (i in 1:n_iter) brute_force_knapsack(
  x = knapsack_objects[1:n_select,], W = 3500, type="matrix"))/n_iter

#- Apply --------------------#
brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="apply1a")
system.time(for (i in 1:n_iter) brute_force_knapsack(
  x = knapsack_objects[1:n_select,], W = 3500, type="apply1a"))/n_iter

brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="apply1b")
system.time(for (i in 1:n_iter) brute_force_knapsack(
  x = knapsack_objects[1:n_select,], W = 3500, type="apply1b"))/n_iter

brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="apply2")
system.time(for (i in 1:n_iter) brute_force_knapsack(
  x = knapsack_objects[1:n_select,], W = 3500, type="apply2"))/n_iter

```
The matrix approach reduces the time with about 2/3 of the time compared to the
for loop, and the combined apply/matrix variant further reduces the time, it 
takes about 1/4 of the time compared to the for loop. The two other apply 
approaches however takes longer time than the for loop.

Another way to improve the running time is to remove the items that have weight
greater than W from the beginning:

```{r}
#- For loop
brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="forloop", 
                     filter_items=TRUE)
system.time(for (i in 1:n_iter) brute_force_knapsack(
  x = knapsack_objects[1:n_select,], W = 3500, type="forloop", filter_items=TRUE)) /n_iter
```
For the data in this example, it takes about 25% of the running time compared to
not filtering the items with weight greater than W.


## 1.1.8 Parallelise brute force search

Three different parallelisation strategies were implemented. The times for the 
different approaches are studied below. Note that it was not possible to set the 
cores to the number of detected cores, since CRAN only allows 2 cores. The results 
for 2 cores are also compared to only using 1 core.

```{r}
#- Set number of knapsack items
n_iter <- 1
n_select <- 18

#- Set number of cores
ncores <- 2

#- Note. number of available cores can be checked with:
ncores_available <- parallel::detectCores()

#- Run algorithms
brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="parallel1a",
                     ncores=ncores)
system.time(for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,],
                                                     W = 3500, type="parallel1a", 
                                                     ncores=ncores))/n_iter

brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="parallel1b", 
                     ncores=ncores)
system.time(for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], 
                                                     W = 3500, type="parallel1b", 
                                                     ncores=ncores))/n_iter

brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="parallel2", 
                     ncores=ncores)
system.time(for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], 
                                                     W = 3500, type="parallel2", 
                                                     ncores=ncores))/n_iter
```
Parallel computations did not improve the running time in these implementations.

Try with only one (1) core:
```{r}
#- Set number of cores
ncores <- 1

#- Run algorithms
brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="parallel1a", 
                     ncores=ncores)
system.time(for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], 
                                                     W = 3500, type="parallel1a", 
                                                     ncores=ncores))/n_iter

brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="parallel1b", 
                     ncores=ncores)
system.time( for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], 
                                                      W = 3500, type="parallel1b", 
                                                      ncores=ncores)) /n_iter

brute_force_knapsack(x = knapsack_objects[1:n_select,], W = 3500, type="parallel2", 
                     ncores=ncores)
system.time( for (i in 1:n_iter) brute_force_knapsack(x = knapsack_objects[1:n_select,], 
                                                      W = 3500, type="parallel2", 
                                                      ncores=ncores)) /n_iter

```
The running times for n=1 core do not differ that much from using n=2 cores in
this implementation.
